---
title: "01 xts time series examples"
author: "John Yuill"
date: "`r Sys.Date()`"
output: 
  html_document:
      code_folding: hide
      toc: true
      toc_depth: 3
      toc_float:
        toc_collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE)

library(tidyverse)
library(lubridate)

library(xts)
library(zoo)
library(quantmod) ## for fetching financial data
library(PerformanceAnalytics)
library(dygraphs)

## theme set for ggplot
theme_set(theme_light())

```

## Purpose

Provide some basic examples of using 'xts' (extensible time series) which is a powerful way to work with time-related data in R.

For more details and examples:

* [xts cheat sheet](https://www.datacamp.com/community/blog/r-xts-cheat-sheet)

## Get some data

* import financial stock data using quantmod pkg, which automatically imports data as 'xts' object

```{r}
getSymbols(Symbols='CSIQ', auto.assign=TRUE)

str(CSIQ)
head(CSIQ)

## drop Adjusted col moving fwd to save space when printing
CSIQ <- CSIQ[,1:5]
```

## Basic filtering/subsetting

### Filter dates

* after given date

```{r}
CSIQ_2020 <- CSIQ['2020-01-01/'] ## from specified date fwd
head(CSIQ_2020)
```

* before given end date

```{r}
CSIQ_2019 <- CSIQ['/2020-01-01'] ## from specified date fwd
tail(CSIQ_2019)
```

* rolling recent period

```{r}
## last n items

CSIQ_7d <- tail(CSIQ, n=7) 
CSIQ_7d
```

* specified period

```{r}
## selected period
CSIQ_Aug <- CSIQ['2020-08/2020-08-31'] ## Aug 1-31
CSIQ_Aug
```

### Filter cols

* as in base R

```{r}
CSIQ_2020_cl <- CSIQ_2020[,4]
head(CSIQ_2020_cl)

plot(CSIQ_2020_cl) ## use basic R plot
```

## Visualize

* several options (where 'x' is your xts object) -> in order shown below
    + plot(x): basic R plotting, will recognize ts and format dates
    + ts.plot(x): specifically for time series, shows dates as index numbers
    + plot.xts(x): very similar to basic plot
    + dygraph(x): often used of financial charts; requires dygraphs pkg
    + chart.TimeSeries(x): from PerformanceAnalytics pkg 

Very simple examples. The last 3 choices above have many additional options, especially for financial data.

```{r}
plot(CSIQ_2020_cl)
ts.plot(CSIQ_2020_cl)
plot.xts(CSIQ_2020_cl)
dygraph(CSIQ_2020_cl)
chart.TimeSeries(CSIQ_2020_cl)
```

## Time Series Manipulation

Leverage xts to work with time series objects 

### End points by period

* day, week, month, qtr, yr.
* returns the locations of each last item in the specified type of period, aligned on calendar.
* will include the last item available in the last period (last data point).
* returns a vector starting with 0 and ending with the extent (last row) of your data.
* can also use k to skip: on='years', k=4 returns location of every 4th yr in data (plus last)

```{r}
## get index number of last day of each year
index_yr <- endpoints(CSIQ, on="years")
## apply to time series to extract selected end points
CSIQ_yr_end <- CSIQ[index_yr]
head(CSIQ_yr_end[,c(2,3,5)]) ## selected cols for space

dygraph(CSIQ_yr_end[,2])

```

```{r}
index_week_end <- endpoints(CSIQ_Aug, on="weeks")
CSIQ_Aug[index_week_end, 1:4] ## open, high, low, close selected
```

### Summarization

```{r}
##
period.apply(CSIQ_Aug, INDEX=index_week_end, FUN=mean)

```

### Shortcut summarization

```{r}
## apply.weekly, apply.monthly, apply.quarterly, apply.yearly to summarize
## FUN=<whatever summarization function for the period)
apply.weekly(CSIQ_Aug, FUN=mean)
```

### Split into periods

```{r}
## split data into weeks - chunk the data into 'week' components, without summarizing
CSIQ_Aug_weeks <- split(CSIQ_Aug, f='weeks')
## second week: high, low, close selected
CSIQ_Aug_weeks[[2]][,2:4]
```

## Aggregate time periods

* Aggregate data by period
* Can take a single variable and calculate OHLC statistics by period

```{r}
CSIQ_yearly <- to.period(CSIQ[,5], period='years', name='Annual', OHLC=TRUE)
CSIQ_yearly
```

## Differences / Change

* Get differences from one period to the next within a time series

```{r}
CSIQ_2020_diff <- diff(CSIQ_2020_cl)
names(CSIQ_2020_diff) <- "Close.diffs"
head(CSIQ_2020_diff)
dygraph(CSIQ_2020_diff)

```

* Differences with lag

```{r}
CSIQ_2020_diff7 <- diff(CSIQ_2020_cl, lag=7)
names(CSIQ_2020_diff7) <- "Close.diff7"
head(CSIQ_2020_diff7, n=10)

CSIQ_diffs <- merge(CSIQ_2020_diff, CSIQ_2020_diff7)
head(CSIQ_diffs)
dygraph(CSIQ_diffs)
```

